Analytic Doppler Imaging

### 2019-08-12 tex/ms.tex
reorg

\newcommand{\sT}{\bvec{s}^\top}
\newcommand{\R}{\bvec{R}}
\newcommand{\AOne}{\bvec{A_1}}
\newcommand{\alm}{\bvec{a}}
\newcommand{\x}{\bvec{x}}
\newcommand{\kT}{\bvec{k}^\top}
\newcommand{\D}{\mathcal{D}}
\newcommand{\Doppler}{\bvec{D}}
\newcommand{\Surf}{\mathcal{S}}
\newcommand{\Curve}{\mathcal{C}}
\newcommand{\Dargs}{\bvec{d}}
\newcommand{\lmax}{\ensuremath{l_\mathrm{max}}}
\newcommand{\vogtstar}{\texttt{VOGTSTAR}\xspace}

% TODO: Use blackboard bold (bbm)
\newcommand{\ylmbasis}{\boldsymbol{\psi}^\top}
\newcommand{\pbasis}{\boldsymbol{\phi}^\top}
\newcommand{\pbasisn}{\ensuremath{\phi_n}}

@@

@@

Finally, we can simplify Equation~(\ref{eq:deconv:F2d}) by noting that the
delta function in the integrand allows us to reduce the double integral 
to a line integral:
%
\begin{align}
    \label{eq:deconv:kT}
    \iint\limits_{\Surf(\x)}
    \ylmbasis(\x)
    \delta\big(\lnlam_0 + \D(\x, \Dargs)\big)
    \mathrm{d}\Surf(\x)
    \, \,
    &=  
    \int\limits_{\Curve(\lnlam, \x, \Dargs)}
    \hspace*{-0.6em}\ylmbasis(\x)
    \mathrm{d}\Curve(\lnlam, \x, \Dargs)
    \nonumber \\[0.5em]
    &\equiv \kT(\lnlam, \Dargs)
    \quad.
\end{align}
%
where the path $\Curve(\lnlam, \x, \Dargs)$ corresponds to the
set of all points on the visible disk where 
$\lnlam_0 + \D(\x, \Dargs) = 0$.
%
We thus have
%
\begin{align}
    \label{eq:deconv:F}
    F(\lnlam, t, \Dargs) 
    &=
    \kT(\lnlam, \Dargs) \, \R(t)
    *
    \alm(\lnlam_0)
    \quad.
\end{align}
%
Equation~(\ref{eq:deconv:F}) is the deconvolution of the
observed spectrum into velocity terms, temporal terms, and spectral/spatial
terms, respectively. In the following sections, we will discuss how to
efficiently solve the line integral in Equation~(\ref{eq:deconv:kT}).

@@

\section{The inverse problem}
\label{sec:inverse}
%
In the previous section we showed how to express the data vector $\bvec{f}$,
a series of spectra obtained at different epochs, as a purely linear
operation on the map vector $\alm$, the spherical harmonic decomposition of
the specific intensity on the surface of the star. The advantage of this
linearity is that, given suitable priors, it allows one to compute both 
the maximum likelihood solution for $\alm$ \emph{and} its uncertainty
\emph{analytically}.

@@

    \sT(\lnlam, \Dargs) 

@@

the vector $\sT(\lnlam, \Dargs)$ in Equation~(\ref{eq:appendix:kT}),

@@

    s_n(\lnlam, \Dargs) 

@@

    s_{i,\,j,\,k}(\lnlam, \Dargs) 

@@

    s_{i,\,j,\,0}(\lnlam_0) 

@@

    s_{0,\,j,\,0}(\lnlam_0) = \frac{j - 1}{j + 1} \big(1 - x^2\big) s_{0,\,j-2,\,0}

@@

    s_{0,\,0,\,0} &= 2 \sqrt{1-x^2} \nonumber \\
    s_{0,\,1,\,0} &= 0 \quad.

@@

    s_{0,\,j,\,1}(\lnlam_0)

@@

    s_{0,\,j,\,1} &= \frac{j - 1}{j + 2} \big(1 - x^2\big) s_{0,\,j-2,\,1}

@@

    s_{0,\,0,\,1} &= \frac{\pi}{2} \big(1-x^2\big) \nonumber \\
    s_{0,\,1,\,1} &= 0 \quad.

@@

    s_{i,\,j,\,k} &= x \, s_{i-1,\,j,\,k} \quad.

@@

    s_{0,\,0,\,0} &= 2 \sqrt{1-x^2} \nonumber \\
    s_{0,\,0,\,1} &= \frac{\pi}{2} \big(1-x^2\big) \nonumber \\
    s_{0,\,1,\,k} &= 0 \quad ,

@@

we may compute any term in $\sT$ via the expressions

@@

    s_{0,\,j,\,k} &= \frac{j - 1}{j + 1 + k} \big(1 - x^2\big) s_{0,\,j-2,\,k} \nonumber \\
    s_{i,\,j,\,k} &= x \, s_{i-1,\,j,\,k} \quad.

>>>>

\newcommand{\R}{\bvec{R}}
\newcommand{\AOne}{\bvec{A_1}}
\newcommand{\alm}{\bvec{a}}
\newcommand{\x}{\bvec{x}}

\newcommand{\D}{D}
\newcommand{\Doppler}{\bvec{D}}
\newcommand{\Surf}{\mathcal{S}}
\newcommand{\Curve}{\mathcal{C}}
\newcommand{\Dargs}{\bvec{d}}
\newcommand{\lmax}{\ensuremath{l_\mathrm{max}}}
\newcommand{\vogtstar}{\texttt{VOGTSTAR}\xspace}

\newcommand{\kT}{\boldsymbol{\kappa}^\top}
\newcommand{\rhoT}{\boldsymbol{\rho}^\top}
\newcommand{\ylmbasis}{\boldsymbol{\psi}^\top}
\newcommand{\pbasis}{\boldsymbol{\phi}^\top}
\newcommand{\pbasisn}{\ensuremath{\phi_n}}

@@

%
\begin{center}
    \begin{longtable}{cll}
    \caption{Notation used in this paper} 
    \label{tab:notation} \\
    %
    \toprule
    \multicolumn{1}{c}{\textbf{Symbol}} &
    \multicolumn{1}{c}{\textbf{Description}} &
    \multicolumn{1}{c}{\textbf{Type}} \\
    \toprule
    \endhead
    %
    \endfoot
    %
    \endlastfoot
    %
    $a$                             & lowercase italic & scalar\\
    $A$                             & uppercase italic & scalar\\
    $\alpha$                        & lowercase italic Greek & scalar\\
    $\bvec{a}$                      & lowercase boldface & column vector\\
    $\bvec{A}$                      & uppercase boldface & matrix\\
    $\boldsymbol{\alpha}^\top$      & lowercase boldface Greek & basis row vector
\end{longtable}
\end{center}

@@

Finally, we can simplify Equation~(\ref{eq:deconv:F2d}) by noting that the
delta function in the integrand allows us to reduce the double integral 
to a line integral:
%
\begin{align}
    \label{eq:deconv:kT}
    \iint\limits_{\Surf(\x)}
    \ylmbasis(\x)
    \delta\big(\lnlam_0 + \D(\x, \Dargs)\big)
    \mathrm{d}\Surf(\x)
    \, \,
    &=  
    \int\limits_{\Curve(\lnlam, \x, \Dargs)}
    \hspace*{-0.6em}\ylmbasis(\x)
    \mathrm{d}\Curve(\lnlam, \x, \Dargs)
    \nonumber \\[0.5em]
    &\equiv \kT(\lnlam, \Dargs)
    \quad.
\end{align}
%
where the path $\Curve(\lnlam, \x, \Dargs)$ corresponds to the
set of all points on the visible disk where 
$\lnlam_0 + \D(\x, \Dargs) = 0$.
%
We thus have
%
\begin{align}
    \label{eq:deconv:F}
    F(\lnlam, t, \Dargs) 
    &=
    \kT(\lnlam, \Dargs) \, \R(t)
    *
    \alm(\lnlam_0)
    \quad.
\end{align}
%
Equation~(\ref{eq:deconv:F}) is the deconvolution of the
observed spectrum into velocity terms, temporal terms, and spectral/spatial
terms, respectively. In the next section, we will discuss how to
efficiently solve the line integral in Equation~(\ref{eq:deconv:kT}).

@@

\section{The inverse problem}
\label{sec:inverse}
%
In the previous section we showed how to express the data vector $\bvec{f}$,
a series of spectra obtained at different epochs, as a purely linear
operation on the map vector $\alm$, the spherical harmonic decomposition of
the specific intensity on the surface of the star. The advantage of this
linearity is that, given suitable priors, it allows one to compute both 
the maximum likelihood solution for $\alm$ \emph{and} its uncertainty
\emph{analytically}. In particular, if one places a (multidimensional)
Gaussian prior on $\alm$ with mean $\boldsymbol{\mu}$ and 
covariance $\boldsymbol{\Lambda}$, the maximum \emph{a posteriori} (MAP) 
solution for $\alm$ is
%
\begin{align}
    \bvec{\hat{a}} &= 
    \boldsymbol{\Sigma}_\mathbf{\hat{a}}
    \left(
        \Doppler^\top
        {\boldsymbol{\Sigma}_\mathbf{f}}^{-1}
        \bvec{f}
        +
        \boldsymbol{\Lambda}^{-1} \boldsymbol{\mu}
    \right)
    \quad,
\end{align}
%
with posterior covariance given by
%
\begin{align}
    \boldsymbol{\Sigma}_\mathbf{\hat{a}} &= 
    \left(
        \Doppler^\top
        {\boldsymbol{\Sigma}_\mathbf{f}}^{-1}
        \Doppler
        +
        \boldsymbol{\Lambda}^{-1}
    \right)^{-1}
    \quad,
\end{align}
%
where $\boldsymbol{\Sigma}_\mathbf{f}$ is the data covariance
matrix. These equations require the inversion of a few fairly
large matrices, but as we will see later, it allows one to obtain
the map of the star ($\bvec{\hat{a}}$) and its uncertainty
($\boldsymbol{\Sigma}_\mathbf{\hat{a}}$) in under a second for
a typical dataset.

There is, however, a catch: for most practical purposes, it is 
difficult to find a good Gaussian prior for $\alm$. Usually
we will have some prior information on what the spectrum of the
star is, and perhaps some prior information on the distribution
of surface features such as starspots. The problem is that
even if these priors are Gaussian, the corresponding prior
on $\alm$ is \emph{not}. To understand this, consider the
matrix $\bvec{A}$, constructed by reshaping $\alm$ into
a matrix of shape $(N \times K')$, whose rows correspond
to the spectral components associated with each spherical harmonic coefficient.
By performing singular value decomposition (SVD) on this matrix, we may
write
%
\begin{align}
    \bvec{A} = \bvec{U} \, \bvec{S} \, \bvec{V}^\top
\end{align}
%
where $\bvec{U}$ is a matrix of shape $(N \times P)$, $\bvec{V}^\top$ is a 
matrix of shape $(P \times K')$, and $\bvec{S}$ is a diagonal matrix of 
shape $(P \times P)$ for some integer $P$.

@@

    \rhoT(\lnlam, \Dargs) 

@@

the vector $\rhoT(\lnlam, \Dargs)$ in Equation~(\ref{eq:appendix:kT}),

@@

    \rho_n(\lnlam, \Dargs) 

@@

    \rho_{i,\,j,\,k}(\lnlam, \Dargs) 

@@

    \rho_{i,\,j,\,0}(\lnlam_0) 

@@

    \rho_{0,\,j,\,0}(\lnlam_0) = \frac{j - 1}{j + 1} \big(1 - x^2\big) \rho_{0,\,j-2,\,0}

@@

    \rho_{0,\,0,\,0} &= 2 \sqrt{1-x^2} \nonumber \\
    \rho_{0,\,1,\,0} &= 0 \quad.

@@

    \rho_{0,\,j,\,1}(\lnlam_0)

@@

    \rho_{0,\,j,\,1} &= \frac{j - 1}{j + 2} \big(1 - x^2\big) \rho_{0,\,j-2,\,1}

@@

    \rho_{0,\,0,\,1} &= \frac{\pi}{2} \big(1-x^2\big) \nonumber \\
    \rho_{0,\,1,\,1} &= 0 \quad.

@@

    \rho_{i,\,j,\,k} &= x \, \rho_{i-1,\,j,\,k} \quad.

@@

    \rho_{0,\,0,\,0} &= 2 \sqrt{1-x^2} \nonumber \\
    \rho_{0,\,0,\,1} &= \frac{\pi}{2} \big(1-x^2\big) \nonumber \\
    \rho_{0,\,1,\,k} &= 0 \quad ,

@@

we may compute any term in $\rhoT$ via the expressions

@@

    \rho_{0,\,j,\,k} &= \frac{j - 1}{j + 1 + k} \big(1 - x^2\big) \rho_{0,\,j-2,\,k} \nonumber \\
    \rho_{i,\,j,\,k} &= x \, \rho_{i-1,\,j,\,k} \quad.
